---
title: "DV_Exercise"
author: "Junxuan Mao"
date: "2/5/2017"
tag:"guns"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Datasource 1: Gunsales - NYTimes
```{r}
## install.packages("gunsales")
library(gunsales)
gun_sales <- analysis()
str(gun_sales[, 1:12])
```

```{r}
plot_gunsales(gun_sales)    
ggplot_gunsales(gun_sales)
```

### Datasource2: Gun Ownership survey data - Washington Post
```{r}
library(XML)
library(dplyr)
gun_ownership_url = 'http://www.washingtonpost.com/wp-srv/health/interactives/guns/ownership.html'
gun_ownership = readHTMLTable(gun_ownership_url, header = TRUE, which = 1)
gun_ownership = gun_ownership[-1, ]

parse_num = function (x) as.numeric(sub(',', '', x))
gun_ownership = select(gun_ownership, State = 1, Total = 2, Yes = 3,
                       `Yes %` = 4, No = 5, `No %` = 6) %>%
                        mutate_each(funs(parse_num), -State)
head(gun_ownership)

gun_ownership = gun_ownership %>%
    mutate(`Child access prevention` = grepl('\\*$', State),
           State  = sub('\\*$', '', State))

# Fix D.C. name
gun_ownership[gun_ownership$State == 'The District', 'State'] = 'District of Columbia' 
```

### Datasource 3: Deaths due to firearms by state - Kaiser Family Foundation
```{r}
gun_deaths = read.csv("/Users/elka/Desktop/Data Visualization/KFF_deaths_firearms.csv", skip = 5) %>%
    select(State = 1, `Deaths per 100000` = 2) %>%
    slice(1:52)

head(gun_deaths)
```

### Dataset 4: Gun-related deaths at the invidual level (2012-2014)
```{r}
library(readr)
gun_inddeaths <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/guns-data/master/full_data.csv")
gun_inddeaths <- gun_inddeaths[,2:dim(gun_inddeaths)[2]]

head(gun_inddeaths)
```


### Questions
1. Does gun ownership correlate with firearm deaths in the U.S.? Does the pattern hold only in the aggregate or also by state?
```{r}
## fix name of "all participant"
gun_ownership[gun_ownership$State == 'All Participants', 'State'] = 'United States'

## joint two data frame by common column "State"
gun_death_onwer <- left_join(gun_deaths, gun_ownership, by = "State")

library(ggplot2)
ggplot(gun_death_onwer, aes(`Yes %`, `Deaths per 100000`)) + geom_point()
```


2. What is the pattern of gunsales over time?
```{r}
library(zoo)
## Combine "year" and "month" into one "Data" column for plotting
gun_sales <- gun_sales %>% mutate(Date = as.yearmon(paste(year, month), "%Y %m"))

ggplot(gun_sales, aes(Date, guns_total)) + geom_point()
ggplot(gun_sales, aes(Date, guns_total_per_1000)) + geom_point()
ggplot(gun_sales, aes(Date, guns_total_per_1000_scaled)) + geom_point()
```


3. What are the individual level patterns of gun deaths? Are gun regulations (e.g. child access prevention laws) associated with lower gun death rates? How about suicides (especially by non-adults)?

```{r}
gun_inddeaths <- gun_inddeaths %>% mutate(Date = as.yearmon(paste(year, month), "%Y %m"))

ggplot(gun_inddeaths, aes(year))+ geom_bar()
ggplot(gun_inddeaths, aes(month))+ geom_bar()
ggplot(gun_inddeaths, aes(Date))+ geom_bar()
ggplot(gun_inddeaths, aes(as.factor(intent)))+ geom_bar()
ggplot(gun_inddeaths, aes(as.factor(sex)))+ geom_bar()
ggplot(gun_inddeaths, aes(age))+ geom_bar()
ggplot(gun_inddeaths, aes(as.factor(race)))+ geom_bar()
ggplot(gun_inddeaths, aes(place))+ geom_bar()
ggplot(gun_inddeaths, aes(education))+ geom_bar()
ggplot(gun_inddeaths, aes(intent, age))+ geom_boxplot()
```



```{r}
## Gun regulations associate wiith lower gun death rates
ggplot(gun_death_onwer, aes(as.factor(`Child access prevention`), `Deaths per 100000`)) + geom_boxplot()
```


