---
title: "Guns Exercise"
author: "Jung Hoon Son"
date: "February 6, 2017"
tag: "guns"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Datasource 1: Gunsales - NYTimes
******

1. Load the __gunsales__ library. 
2. __analysis()__ (not sure what it does). 
3. Display structure of **gun_sales**.

```{r}
library(gunsales)
gun_sales <- analysis()
str(gun_sales[,1:12])
```
******
```{r}
plot_gunsales(gun_sales)    
ggplot_gunsales(gun_sales)
```
   
******
## Datasource 2: Gun ownership survey data - Washington Post
******


```{r}
library(XML)
library(dplyr)
gun_ownership_url = 'http://www.washingtonpost.com/wp-srv/health/interactives/guns/ownership.html'
gun_ownership = readHTMLTable(gun_ownership_url, header = TRUE, which = 1)
gun_ownership = gun_ownership[-1, ]

parse_num = function (x) as.numeric(sub(',', '', x))
gun_ownership = select(gun_ownership, State = 1, Total = 2, Yes = 3,
                       `Yes %` = 4, No = 5, `No %` = 6) %>%
                        mutate_each(funs(parse_num), -State)
head(gun_ownership)
```
   
******
## Datasource 2: Gun ownership survey data - Washington Post
******

```{r}
gun_ownership = gun_ownership %>%
    mutate(`Child access prevention` = grepl('\\*$', State),
           State  = sub('\\*$', '', State))

# Fix D.C. name
gun_ownership[gun_ownership$State == 'The District', 'State'] = 'District of Columbia'
```
   
******
## Datasource 3: Deaths due to firearms by state - Kaiser Family Foundation
******
```{r}
# Website appears to block scraping, but allows downloading data.
# gun_deaths_url = 'http://kff.org/other/state-indicator/firearms-death-rate-per-100000/'
# gun_deaths = readHTMLTable(gun_deaths_url)
# Instead, we use the manually downloaded CSV output.
gun_deaths = read.csv("exercise/KFF_deaths_firearms.csv", skip = 3) %>%
    select(State = 1, `Deaths per 100000` = 2)
head(gun_deaths)
```
   
******
## Dataset 4: Gun-related deaths at the individual level (2012-2014)
******
```{r}
library(readr)
gun_inddeaths <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/guns-data/master/full_data.csv")
gun_inddeaths <- gun_inddeaths[,2:dim(gun_inddeaths)[2]]
head(gun_inddeaths)
```
   
***
## Possible Questions to explore with these data (feel free to explore what you think is interesting)
***
## Does gun ownership correlate with firearm deaths in the U.S.? Does the pattern hold only in the aggregate or also by state?

After looking at data, rows 2 to 52 show all 50 states+DC for both **gun_deaths** and **gun_ownership** tables, in alphabetical order. Created a single data table **deaths_ownership** for easier plotting/management.

```{r}
deaths_ownership<-cbind(gun_deaths[2:52,],ownership= gun_ownership[2:52,]$'Yes %')
colnames(deaths_ownership)[2]<-"death_rate"
library(ggplot2)
ggplot(deaths_ownership,aes(ownership,death_rate,label=State))+
  geom_point(data=deaths_ownership) +
  geom_text(size=3,
            vjust = 0,
            nudge_y = 0.2, 
            nudge_x = 0.1) +
  labs(title="Deaths Rate vs. Gun Ownership", 
       y="Deaths per 100000",
       x="Gun Ownership")
```
   
## What is the pattern of gunsales over time?
***
```{r}
library(scales)
# A bit messy way to format time
gun_sales$formatted_date<-as.Date(strptime(paste0(gun_sales$year,'-',gun_sales$month,'-01'),format="%Y-%m-%d"))
# Plot with scale_x_date, which prevents ggplot from trying to display all the dates
ggplot(gun_sales,
       aes(formatted_date,guns_total_per_1000)) +
  geom_point(data=gun_sales) +
  labs(title="Gun sales over time",
       y="Gun Sales per 1000",
       x="Date") +
  scale_x_date(labels = date_format("%Y-%m"))
```
   
## What are the individual level patterns of gun deaths? Are gun regulations (e.g. child access prevention laws) associated with lower gun death rates? How about suicides (especially by non-adults)?
******
```{r}
# A bit messy way to format time
gun_inddeaths$formatted_date<-as.Date(strptime(paste0(gun_inddeaths$year,'-',gun_inddeaths$month,'-01'),format="%Y-%m-%d"))
```
   
### Gun Deahts, by Intent
******
```{r}
gun_inddeaths %>% 
  na.omit() %>% 
  ggplot(aes(intent,fill=intent)) + 
  geom_bar(stat="count")+
  theme( legend.position="none")+
  labs(title="Gun deaths, by intent")
```
   
### Gun Deaths, by Race
******
```{r}
gun_inddeaths %>%
  na.omit() %>%
  ggplot(aes(race,fill=race)) + 
  geom_bar(stat="count")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1), legend.position="none")+
  labs(title="Gun deaths, by Race")
```
   
### Gun Deaths, by Place
******
```{r}
gun_inddeaths %>%
  na.omit() %>%
  ggplot(aes(place,fill=place)) + 
  geom_bar(stat="count")+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1), legend.position="none")+
  labs(title="Gun deaths, by Place")

```
   
### Gun regulation and death rates among states
******
```{r}
deaths_regulations<-cbind(gun_deaths[2:52,],child_regulation=gun_ownership[2:52,c(7)])
colnames(deaths_regulations)[2]<-"death_rate"

ggplot(deaths_regulations,aes(child_regulation,death_rate,colour=child_regulation)) + 
  geom_boxplot() +
  labs(title="Gun regulation and death rates",
       y="Deaths per 100,000",
       x="Child Regulation in state")+
  theme( legend.position="none")+
  scale_x_discrete(labels=c("No","Yes"))
```

### Suicides by non-adults (age<18)
******
```{r}
gun_inddeaths %>%
  filter(intent=="Suicide") %>%
  filter(age<18) %>%
  group_by(date=formatted_date) %>%
  summarize(counts=n()) %>%
  ggplot(aes(x=date,
           y=counts)) + 
  geom_line() +
  labs(title="Gun Suicides of Non-adults over time (2012-2015)",
       y="Deaths",
       x="Date")
```